---
title: "Lab 2 Homework"
author: "Maxen Hamelynck"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

## Instructions

Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!

1.  What is a vector in R?

Vectors are methods of organizing data into objects as sequences, strings, or simple lists

2.  What is a data matrix in R?

Matrices are stacks of vectors organized such that they appear as data tables; they're another method of organizing data.

3.  Below are data collected by three scientists (Jill, Steve, Susan in order) measuring temperatures of eight hot springs. Run this code chunk to create the vectors.\

```{r}
spring_1 <- c(36.25, 35.40, 35.30)
spring_2 <- c(35.15, 35.35, 33.35)
spring_3 <- c(30.70, 29.65, 29.20)
spring_4 <- c(39.70, 40.05, 38.65)
spring_5 <- c(31.85, 31.40, 29.30)
spring_6 <- c(30.20, 30.65, 29.75)
spring_7 <- c(32.90, 32.50, 32.80)
spring_8 <- c(36.80, 36.45, 33.15)
```

4.  Build a data matrix that has the springs as rows and the columns as scientists.

```{r}
temperatures <- c(spring_1, spring_2, spring_3, spring_4, spring_5, spring_6, spring_7, spring_8)
people <- c('Jill', 'Steve', 'Susan')
hotspring_matrix <- matrix(temperatures, nrow=8, byrow=T)
colnames(hotspring_matrix) <- people
hotspring_matrix
```

4.  The names of the springs are 1.Bluebell Spring, 2.Opal Spring, 3.Riverside Spring, 4.Too Hot Spring, 5.Mystery Spring, 6.Emerald Spring, 7.Black Spring, 8.Pearl Spring. Name the rows and columns in the data matrix. Start by making two new vectors with the names, then use `colnames()` and `rownames()` to name the columns and rows.

    ```{r}
    people <- c('Jill', 'Steve', 'Susan')
    colnames(hotspring_matrix) <- people

    springnames <-c('Bluebell Spring','Opal Spring','Riverside Spring', 'Too Hot Spring','Mystery Spring','Emerald Spring', 'Black Spring', 'Pearl Spring')
    rownames(hotspring_matrix) <- springnames

    hotspring_matrix
    ```

5.  Calculate the mean temperature of all eight springs.

    ```{r}
    mBluebell <- mean(hotspring_matrix[1:1,1:3])
    mOpal <- mean(hotspring_matrix[2:2, 1:3])
    mRiver <- mean(hotspring_matrix[3:3, 1:3])
    mHot <- mean(hotspring_matrix[4:4, 1:3])
    mMyster <- mean(hotspring_matrix[5:5,1:3])
    mEmerald <- mean(hotspring_matrix[6:6,1:3])
    mBlack <- mean(hotspring_matrix[7:7,1:3])
    mPearl <- mean(hotspring_matrix[8:8,1:3])

    mBluebell 
    mOpal 
    mRiver 
    mHot 
    mMyster 
    mEmerald 
    mBlack 
    mPearl 

    rowSums(hotspring_matrix)/3 

    ```

6.  Add this as a new column in the data matrix.

    ```{r}
    Average_temperatures <- rowSums(hotspring_matrix)/3 
    averagehotspring_matrix <- cbind(hotspring_matrix, Average_temperatures)
    averagehotspring_matrix
    ```

7.  Show Susan's value for Opal Spring only.

    ```{r}
    mean(hotspring_matrix[,3])

    ```

8.  Calculate the mean for Jill's column only.

    ```{r}
    mean(hotspring_matrix[,1])

    ```

9.  Use the data matrix to perform one calculation or operation of your interest.

    Calculating the integer mean of Steve's hotspring data from 'Too hot' to 'Pearl'

    And reexpressing the matrix only as integers greater than or equal to 31

    ```{r}


    partialcol2 <- hotspring_matrix[4:8, 2]
    partialcol2

    intpartialcol2 <- as.integer(partialcol2)
    intpartialcol2

    mean(intpartialcol2)



    inttemp <- as.logical(temperatures>=31)
    intmatrix <- matrix(inttemp, nrow=8, byrow=T)
    intmatrix
    ```

## Push your final code to GitHub!

Please be sure that you check the `keep md` file in the knit preferences.
